---
title: "Importing Panel Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Importing Panel Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#>devtools::install_github("antaldaniel/eurobarometer")
library(eurobarometer)
```

The files are placed in the `not_included` directory.  To reproduce this example, you have to place exactly the same GESIS files in the `not_included` subfolder of your project, or modify the paths and file names according to your project.

The SPSS file is imported with `haven`, which is part of the `tidyverse`.


```{r imports}
require(haven)
ZA5479_raw <- haven::read_spss(file.path("..", "not_included", "ZA5479_v6-0-0.sav"))
```

The `gesis_metadata_create` and `canonical_name_create` will create suggested canonical names from the GESIS file, and various other pre-processing steps.

```{r metadata}
ZA5479_metadata <- gesis_metadata_create(dat = ZA5479_raw)
```

```{r renaming}
ZA5479
```
