---
title: "Working With Vocabularies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vocabulary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE, 
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eurobarometer)
library(dplyr)
library(tibble)
library(knitr)
library(kableExtra)
# The examples of this vignette can be found run with  
# source(
#  file.path("not_included", "vignette_vocabulary_examples.R")
#  )
```

There is no final metadata solution now.  There are two main functions, [gesis_metadata_create](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html), which mainly helps with normalizing the variable names, and [gesis_vocabulary_create](http://eurobarometer.danielantal.eu/reference/gesis_vocabulary_create.html) (wrapped by [gesis_corpus_create](http://eurobarometer.danielantal.eu/reference/gesis_corpus_create.html)), which help consolidating the vocabulary of questionnaire items.

They are both referring back to [normalize_names](http://eurobarometer.danielantal.eu/reference/normalize_names.html) in normalizing labels. The  [canonical_names](http://eurobarometer.danielantal.eu/reference/canonical_names.html) is a very early attempt to use a canonical form of same questions that appear under different normalized forms.

This is a database created from 28 Eurobarometer SPSS files.

```{r}
metadata_database <-  readRDS( 
  file.path('..', 'data-raw',
  'eb_metadata_database.rds')) 
```

Let's filter out a typical question type that has `Tend to trust` and `Tend not to trust` answer options, with differently encoded `Declined` to answer options.

```{r}
select_metadata_vars <- c("filename", "r_name", "normalized_names", 
                          "canonical_names",
                          "label_normalized" , "label",
                          "item_no", "item_of")

trust_metadata <- metadata_database  %>%
  filter (
    grepl( "tend_to_trust|tend_not_to_trust", label_normalized )
  )  %>%
  select ( all_of(select_metadata_vars)) %>%
  arrange ( normalized_names, label_normalized, filename )
```

To get the unique questions, let's leave out the `Tend not to trust` negative answer options. 

```{r}
require(kableExtra)
trust_metadata %>% filter ( 
   grepl( "tend_to_trust", label_normalized )
  ) %>%
  sample_n(15) %>% # Print only a sample of 15 rows
kable() %>%
  kable_styling(bootstrap_options =
                c("striped", "hover", "condensed"), 
                fixed_thead = T, 
                font_size = 7)
```

Let's add back all value labels:

```{r}
trust_metadata %>%
  select (
    # previously we filtered tend to (not) agree, remove it
    -all_of(c("label_normalized", "label"))
    ) %>%
  left_join ( 
    # add back all value labels, i.e. various declines
    metadata_database %>%
                select ( all_of(select_metadata_vars))
    )  %>% 
  sample_n(15) %>% # Print only a sample of 15 rows
kable() %>%
  kable_styling(bootstrap_options = 
                c("striped", "hover", "condensed"), 
                fixed_thead = T, 
                font_size = 9)
```

## Harmonization of Value Labels

Let's review the problems here. 

```{r, results='asis'}
trust_with_decline  <- metadata_database %>%
  select ( all_of(select_metadata_vars) ) %>%
  semi_join ( trust_metadata %>%
                select ( all_of(c("filename", "r_name"))), 
              by = c("filename", "r_name"))

trust_var_labels <- trust_with_decline %>%
  select ( all_of(c("label", "label_normalized", "item_no", "item_of"))) %>%
  distinct_all () %>%
  arrange ( label_normalized, label, item_of)

trust_var_labels %>% 
kable() %>%
  kable_styling(bootstrap_options = 
                c("striped", "hover", "condensed"), 
                fixed_thead = T, 
                font_size = 9 ) 
```

We are dealing with different scales here. Let's narrow down. 

Some variables appear to have four categories, because the declined answers are differently coded in the Turkish Community of Cyprus.  In other cases, though, this is just a different scale.

```{r}
trust_2_with_decline  <- trust_with_decline %>%
  filter ( item_of <= 4)

trust_2_value_labels <- trust_2_with_decline %>%
  select ( all_of(c("label", "label_normalized", 
                    "item_no", "item_of"))
           ) %>%
  distinct_all () %>%
  arrange ( label_normalized, label, item_of)

trust_2_value_labels %>% 
kable() %>%
  kable_styling(bootstrap_options = 
                c("striped", "hover", "condensed"), 
                fixed_thead = T, 
                font_size = 10 ) 

```

My approach would be this:
1. filter out all questions with normalized answer options `tend_to_trust` and `tend_not_to_trust`.  These can be safely harmonized.

2. There is clearly a lot of work to be done with the refusal options. I think that this would merit a different function and a different vocabulary, because missingness should consistently treated anyway. The treatment of missinginess in the R language will also raise issues, because the `NA` values is not class independent.

3. Agree on a class representation of the value labels and the missingness. A binary variable would lend itself to be numerically represented, but if we want to treat trust as a topic, which is sometimes coded differently, then a factor or character representation is better (and we should probably use their more modern vctrs versions.).

4. A factor representation is useful if we want to use the ordinary scale of the variable. In this case we must make sure that at each table join, the number and ordering of the factor levels is consistent.

5. I tend to belive that it is better to add back the ordinary scaling later, becasue it is not always necessary at all. (For example, the researcher may want to use this variables in PCA or select a few of them as indicators/dummies.) It is far more simple to give a character representation to these labels. 

6. Manually find out what is going on with 	`tend_to_trust_7_8_in_qa8` `tend_not_to_trust_1_4_in_qa8`. If they can be safely treated as an exception and recoded to `tend_to_trust` and `tend_not_to_trust`, add them to the controlled vocabulary.

7. Leave out anything that has different labels than `tend_to_trust`,  `tend_not_to_trust` or their missingness value.

```{r}
trust_vocabulary <- tibble::tibble (
  # maybe we can use a generic controlled vocabulary, for example
  # Library of Congress
  topic_1 = 'trust',
  # And if we find them, we can add GESIS or TNS/Kantar keywords here
  topic_2 = 'trust, binary',
  label_normalized = trust_2_value_labels %>%
    filter ( 
      grepl("tend_to|tend_not_to|inap|dk", label_normalized)) %>%
    distinct ( label_normalized ) %>%
    unlist () %>%
    as.character(), 
  level = 3 # missingness should be harmonized in character form 
) 

trust_table <- trust_vocabulary %>%
  mutate( 
    character_value = case_when(
      # and create a surely harmonized character representation 
      grepl("dk|inap", label_normalized) ~ NA_character_, 
      substr(label_normalized, 1,7) == "tend_to" ~ "tend_to_trust", 
    TRUE ~ "tend_not_to_trust"), 
    numeric_value = case_when ( 
      character_value == "tend_to" ~ 1,
      character_value == "tend_not_to" ~ 0,
      TRUE ~ NA_real_),
    missing = case_when (
      # it is useful for faster filtering of missingness
      # and true value labels  
      is.na(numeric_value) ~ TRUE, 
      TRUE ~ FALSE)
  )

trust_table %>%
  kable %>%
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed"), 
                  fixed_thead = T, 
                  font_size = 10 ) %>%
    add_header_above(c("Keywords" = 2, 
                       "Label Identification" = 2, 
                       "Value Harmonization" = 3))
```

As far as I see, this can be a bulding block of an important metadata table, for example, a table of two-level categories.  As far as I see it, this table is capable of harmonizing the two-level trust variables in all 28 Eurobarometers I checked.

The declined answers should be consistently added to the levels. Currently I used `NA_character_` in the character representation, but I think that a harmonized character version, such as the often used `declined` would be better. This could consistently treated as `NA_real_` in a suggested numerical representation, whenever that is applicable.


```{r}
saveRDS(trust_table, file.path(
  "..",   # we're in vignettes
  "data-raw", "trust_value_labels.rds"),
  version = 2 # downward compatibility on CRAN
  )
```

Once we have a few more bits, for example, similar elements of other 2-value factors, we'll bind them together in the R file `data-raw/create_categorical_var.R`.

Only the dataset maintainer should run `usethis::use_data()`, which creates the final file `data/category_labels_2`. 

When the data file is updated, you should check the documentation file `R/data-category_label_2.R`. This will be used to automatically update the data documentation.  At last, `devtools::document()` will automatically update the `man/data-category_label_2.Rd` manual file.  Never change the contents of the `man` directory by hand!

The data documentation from the manual file can be reviewed by `?category_labels_2`.

The `category_labels_2` should be extended by similar, two-level categorical variables later.  The current stage can be reviewed with `data(category_labels_2`).

## Harmonization of Variable Names

Please review [Harmonizing Variable Names](http://eurobarometer.danielantal.eu/articles/harmonization.html) vignette.  Pull requests are welcome: if you want to make changes, or add details, please, modify the `variable_names.Rmd` file, and not the `.Rd` or `.html` or other compiled files.  If you are unsure that we are woking paralell, just make a copy of the file to `not_included` with your version.

```{r}
trust_variable_table <- trust_with_decline %>%
  filter ( label_normalized %in% trust_table$label_normalized ) %>%
  filter ( ! grepl("_recoded", normalized_names ) ) %>%
  distinct ( label_normalized, .keep_all = TRUE ) %>%
  mutate ( institution =      gsub("trust_in_institutions_|trust_in_|_trust", "",
                                   normalized_names )) %>%
  mutate ( geo_qualifier = case_when(
    grepl("_tcc", institution) ~ "tcc",
    TRUE ~ NA_character_),
    institution = gsub("_tcc", "", institution)
  ) %>%
  mutate ( canonical_names = paste0("trust_in_",
                                    institution, "_",
                                    geo_qualifier) ) %>%
  mutate ( canonical_names = gsub("_NA", "", canonical_names)) %>%
  select ( -contains("label"),
           -contains("item")) %>%
  rename ( keyword_1 = institution )

trust_variable_table %>%
  kable %>%
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed"), 
                  fixed_thead = T, 
                  font_size = 10 ) %>%
    add_header_above(c("Filtering" = 3, 
                       "Preferred Term" = 1, 
                       "Keywords" = 2)
                     )
```

I think that this can be the basis of a metadata table to use canonical names for 2-level factors.

We should make an effort to harmonize the keywords(s) with accepted thesauri, in this case, for example, with the [ELSST](https://elsst.ukdataservice.ac.uk/thesaurus-search/view-concept/?id=78e0b701-e501-48a1-864f-2d74fb12a37c&lang=EN).

```{r}
saveRDS(trust_variable_table, file.path(
  "..",   # we're in vignettes
  "data-raw", "trust_variables.rds"),
  version = 2 # downward compatibility on CRAN
  )
```

Once we have a few more bits, for example, similar elements of other 2-value factors, we'll bind them together in the R file `data-raw/create_categorical_var`. [Same as the previous data file, but the data files will be separate.]

Only the dataset maintainer should run `usethis::use_data()`, which creates the final file `data/categorical_variables_2`. 

When the data file is updated, you should check the documentation file `R/data-categorical_variables_2.R`. This will be used to automatically update the data documentation.  Like in the previous example, with `devtools::document()` will automatically update the `man/data-categorical_variables_2.Rd` manual file.  Never change the contents of the `man` directory by hand!

The data documentation from the manual file can be reviewed by `?categorical_variables_2`.

The `categorical_variables_2` should be extended by similar, two-level categorical variables later.  The current stage can be reviewed with `data(categorical_variables_2`).

## Renaming Suggestions

If you are unhappy with any names used in this package itself (see all functions and arguments on the  [reference](http://eurobarometer.danielantal.eu/reference/index.html)page), please, let me know in an email.  We should agree on all column names of metadata files, function names and function parameters as early as possible, because it will be extremely difficult to make name changes later as more and more code piles up.  
